---
title: "INLA scale code study"
author: "Hashan Fernando"
date: "2024-03-06"
output:
  html_document: default
  pdf_document: default
---

```{r}
library(INLA)   
#df <- read.csv("Qmatrix_EP_DISABL.csv", header = TRUE)



#df <- as.matrix(read.csv("Qmatrix_EP_CROWD.csv", header = TRUE))
#mat_disabl <- as.matrix(read.csv("Qmatrix_EP_DISABL.csv", header = TRUE))
#df <- as.matrix(read.csv("Qmatrix_EP_LIMENG.csv", header = TRUE))
#mat_pci <- as.matrix(read.csv("Qmatrix_EP_PCI.csv", header = TRUE))
df <- as.matrix(read.csv("Qmatrix_EP_UNINSUR.csv", header = TRUE))


mat <- as.matrix(df)
mat
```
In constr \

A: It specifies a coefficient matrix with dimensions 1x12 (1 row and 12 columns), filled with ones. This implies that the linear combination of the 12 variables in the model should sum to a constant value. \

e: It specifies a constant value of 0 for the right-hand side of the linear constraints. This means that the linear combination of variables should equal zero.


inla.scale.model Study
\


define 12 element vector of 0 \ 
rep() function is used to replicate elements of vectors or array \

```{r}
marg.var <- rep(0, nrow(mat))
marg.var
```

inla.as.sparse(Q) same as dgTMatrix \

dgTMatrix format: The dgTMatrix format is a compressed representation of a sparse matrix, particularly efficient for matrices with a large number of zero elements. It stores only the non-zero elements and their positions, making it memory-efficient and suitable for various mathematical operations \


```{r}
Q_sparse <- inla.as.sparse(mat)
Q_sparse
```
```{r}
g <- inla.read.graph(Q_sparse)
```

```{r}
constr=list(A=matrix(1, nrow=1, ncol=nrow(mat)),e=0)
eps = sqrt(.Machine$double.eps)
```

```{r}
for (k in seq_len(g$cc$n)) {
        i <- g$cc$nodes[[k]]
        n <- length(i)
        QQ <- mat[i, i, drop = FALSE]
        if (n == 1) {
            QQ[1, 1] <- 1
            marg.var[i] <- 1
        }
        else {
            cconstr <- constr
            if (!is.null(constr)) {
                cconstr$A <- constr$A[, i, drop = FALSE]
                eeps <- eps
            }
            else {
                eeps <- 0
            }
            idx.zero <- which(rowSums(abs(cconstr$A)) == 0)
            if (length(idx.zero) > 0) {
                cconstr$A <- cconstr$A[-idx.zero, , drop = FALSE]
                cconstr$e <- cconstr$e[-idx.zero]
            }
            #L <-QQ + Diagonal(n)*max(diag(QQ))*eeps
            
          #print(QQ)
            
            res <- inla.qinv(QQ + Diagonal(n) * max(diag(QQ)) * 
                eeps, constr = cconstr)
            
            fac <- exp(mean(log(diag(res))))
            print(fac)
            QQ <- fac * QQ
            marg.var[i] <- diag(res)/fac
        }
        
}
```


```{r}
sum(0.2420615,0.4550647,0.25)/3
```


















```{r}
plot(g)
summary(g)
```


INLA graph function \

We need g$cc$n \
and  g$cc$nodes \


```{r}
print(g$cc$n)
print(g$cc$nodes )
```



```{r}
i <- g$cc$nodes[[1]]
print(i)
n <- length(i)


# This indicates that we are selecting rows and columns from the matrix specified by the indices i. Since i is used for both rows and columns,
QQ <- Q_sparse[i, i, drop = FALSE]


if (n == 1) {
  QQ[1, 1] <- 1
  marg.var[i] <- 1
}

```
```{r}
QQ
```

Else part(important)
```{r}
L =list(A=matrix(1, nrow=1, ncol=12))
L
```
constr	\
Linear constraints spanning the null-space of Q; see ?INLA::f and argument extraconstr \

When we talk about linear constraints spanning the null-space of the Q matrix, it means that we are imposing linear restrictions on the variables represented by x such that the resulting vector lies in the null-space of Q. In other words, these constraints are defining relationships between the variables such that the resulting combination satisfies the equation Qx=0. \

For example, if Q is a coefficient matrix representing a system of linear equations, the null-space contains all the solutions to the homogeneous system Qx=0. In this context, linear constraints spanning the null-space could refer to additional conditions imposed on the variables to ensure that they satisfy the given system of equations. These constraints would help define the solution space more precisely. \

eps	
A small constant added to the diagonal of Q if constr \

If there constrain eeps will be Machine$double.eps else it will be zero.

```{r}
constr = L
eps = sqrt(.Machine$double.eps)

cconstr <- constr

if (!is.null(constr)) {
    cconstr$A <- constr$A[, i, drop = FALSE]
    eeps <- eps
}else {
    eeps <- 0
}

```

```{r}
constr$A[, i, drop = FALSE]
```

```{r}
constr$A

```



The which() function in R returns the indices of the elements that satisfy a given condition. In this specific case, which(rowSums(abs(cconstr$A)) == 0) will return the indices of the rows in the matrix cconstr$A where the sum of the absolute values of elements across each row equals zero.

```{r}
idx.zero <- which(rowSums(abs(cconstr$A)) == 0)
print(idx.zero)
```
```{r}
if (length(idx.zero) > 0) {
    #This line removes the rows indexed by idx.zero from the matrix cconstr$A. The -idx.zero notation is used to index all rows except those specified in idx.zero. The drop = FALSE argument ensures that the result remains a matrix even if it has only one row.
    cconstr$A <- cconstr$A[-idx.zero, , drop = FALSE]
    #  This line removes the elements indexed by idx.zero from the vector cconstr$e
    cconstr$e <- cconstr$e[-idx.zero]
}
```


```{r}
Diagonal(n)
```
```{r}
max(diag(QQ))
```


```{r}
Diagonal(n)*max(diag(QQ))*eeps
```
```{r}
QQ+Diagonal(n)*max(diag(QQ))*eeps
```

Have to define this as 0. Look the function e=0
```{r}
cconstr$e <-0
```

```{r}
res <- inla.qinv(QQ + Diagonal(n)*max(diag(QQ))*eeps,constr = cconstr)
res
```
```{r}
fac <- exp(mean(log(diag(res))))
fac
```


```{r}
QQ_ <- fac * QQ
QQ_
```








```{r}
Q_sparse[c(1,2), c(1,2), drop = FALSE]
```



```{r}
marg.var[i] <- diag(res)/fac
marg.var
```

```{r}
Q
```

```{r}
QQ_
```
Putting it all together, Q[i, i] <- QQ means that the elements of the square submatrix QQ are assigned to the diagonal elements of the matrix Q at positions specified by the indices i. This operation effectively updates the diagonal elements of Q with the values from QQ.
 
 
 
 
```{r}
# Check the number of rows
num_rows <- nrow(Q)

# Check the number of columns
num_cols <- ncol(Q)

# Print the number of rows and columns
print(num_rows)
print(num_cols)
```

```{r}
# Check the number of rows
num_rows <- nrow(QQ_)

# Check the number of columns
num_cols <- ncol(QQ_)

# Print the number of rows and columns
print(num_rows)
print(num_cols)
```



```{r}
#Q[i, i] <- QQ_
```




















```{r}
my_function <- function (Q, constr = NULL, eps = sqrt(.Machine$double.eps)) 
{
    marg.var <- rep(0, nrow(Q))
    Q <- inla.as.sparse(Q)
    g <- inla.read.graph(Q)
    for (k in seq_len(g$cc$n)) {
        i <- g$cc$nodes[[k]]
        n <- length(i)
        QQ <- Q[i, i, drop = FALSE]
        if (n == 1) {
            QQ[1, 1] <- 1
            marg.var[i] <- 1
        }
        else {
            cconstr <- constr
            if (!is.null(constr)) {
                cconstr$A <- constr$A[, i, drop = FALSE]
                eeps <- eps
            }
            else {
                eeps <- 0
            }
            idx.zero <- which(rowSums(abs(cconstr$A)) == 0)
            if (length(idx.zero) > 0) {
                cconstr$A <- cconstr$A[-idx.zero, , drop = FALSE]
                cconstr$e <- cconstr$e[-idx.zero]
            }
            res <- inla.qinv(QQ + Diagonal(n) * max(diag(QQ)) * 
                eeps, constr = cconstr)
            fac <- exp(mean(log(diag(res))))
            QQ <- fac * QQ
            marg.var[i] <- diag(res)/fac
        }
        
        #print(Q)
        #print(QQ)
        Q[i, i] <- QQ
        print(Q)
    }
    return(list(Q = Q, var = marg.var))
}
```

```{r}
T = my_function(mat, constr=list(A=matrix(1, nrow=1, ncol=12),e=0))
```
```{r}
TT = T$Q
```

```{r}
T
```


```{r}
help(inla.qinv)
```

SPD stands for "Symmetric Positive Definite." An SPD sparse matrix is a specific type of sparse matrix that possesses two key properties:

Symmetry: The matrix is symmetric, meaning that it is equal to its transpose. Mathematically, for an × n×n matrix A, this property is represented as A=A⊤ . This symmetry often arises in many applications, particularly in contexts where relationships or interactions are symmetric.

Positive Definiteness: The matrix is positive definite, meaning that it satisfies certain mathematical properties related to its eigenvalues. Specifically, all eigenvalues of the matrix are positive. In other words, for any non-zero vector x, the quadratic form x⊤ Ax is always positive.

```{r}
getAnywhere(inla.qinv)
```
```{r}
d = inla.scale.model(mat)



